---
title: "HW2_rka"
author: "Radhakrishna Adhikari"
date: "12/10/2021"
output: pdf_document
---

# Homework 2

## Problem 2

### Part A

Since I am submitting this homework at the end of semester, I am writing what I learned in this class. Even though I had been using other programming languages like Python, Matlab, and Java for several years, I learned R and used R for the first time in this class. I found the book recommended for the class, "R for data science" very helpful. I also learned to write the R Markdown files, and learned to write using LaTex. Overall, it was really helpful class to learn the R.

### Part B
I have listed some distribution function as following.

Lognormal Density Function:

\begin{align}
f\left(x|\mu,\sigma^2\right)=\frac{1}{\sqrt{2\pi}\sigma}\frac{e^{-\left(\log x-\mu\right)^2/\left(2\sigma^2\right)}}{x}; 0\leq x<\infty; -\infty<\mu<\infty
\end{align}

Gamma Density Function:

\begin{align}
f \left(x| \alpha,\beta \right)= \frac{1}{ \Gamma \left( \alpha \right) \beta^ \alpha}x^{ \alpha-1}e^{-x/ \beta}; 0 \leq x< \infty; \alpha,\beta > 0
\end{align}

Chi squared Density Function:

\begin{align}
f \left(x|p \right)= \frac{1}{ \Gamma \left(p/2\right) 2^{p/2}}x^{\left(p/2\right)-1}e^{-x/2}; 0\leq x<\infty ;p=1,2,... 
\end{align}


## Problem 3

\begin{enumerate}
\item Track steps to record how all results were produced
\begin{itemize}
\item Challenges: These results are often produced through a lot of trial and error , so it can be difficult to separate the steps that were important and those that were unnecessary.
\end{itemize}
\item Do not manually manipulate data
\begin{itemize}
\item Challenges: Some machines may not be able to read specific data types, but opening the file with a different type may change the data without your knowledge.
\end{itemize}
\item Archive (or keep track of) the exact versions of all programs used
\begin{itemize}
\item Challenges: Even if you do archive the program, some updates will automatically delete older versions on the computer, possibly rendering past research useless.
\end{itemize}
\item Version control all scripts(use github/bitbucket)
\begin{itemize}
\item Challenges: It can be difficult to know when to store versions of code and when not to. Too many versions of code can still make the correct version difficult to find, and too few means you are less likely to have the exact code that you want. 
\end{itemize}
\item Record all intermediate results and standardize if possible
\begin{itemize}
\item Challenges: Intermediate steps might not be in a data form that is easy to save, so it might not be possible to keep track of all intermediate steps.
\end{itemize}
\item Note seeds used for analyses that include randomness
\begin{itemize}
\item Challenges: Using seeds for randomness may not be appropriate in the context of the experiment.
\end{itemize}
\item Store raw data used to make plots
\begin{itemize}
\item Challenges: Large data sets may require more storage space than what is available.
\end{itemize}
\item Keep and inspect all layers of detail of the data
\begin{itemize}
\item Challenges: Amount of data to inspect can grow quickly if there are a lot of layers of data.
\end{itemize}
\item Connect statements to the results that inspired them
\begin{itemize}
\item Challenges: Research in a specialized field can be difficult to explain to the general public.
\end{itemize}
\item Provide public access to all data used, programs written, and results discovered
\begin{itemize}
\item Challenges: Some data is not publicly available and perhaps must be purchased, so it cannot be included with the paper.
\end{itemize}
\end{enumerate}

## Problem 4

``` {r}
#install.packages('data.table')
library(data.table)
covid_raw = fread("https://opendata.ecdc.europa.eu/covid19/casedistribution/csv")
us = covid_raw[covid_raw$countriesAndTerritories == 'United_States_of_America',]
us_filtered = us[us$month %in% c(6:7),]
us_filtered$index = rev(1:dim(us_filtered)[1]) 
fit=lm(`Cumulative_number_for_14_days_of_COVID-19_cases_per_100000`~index, data=us_filtered)
```

### Part A

#### Part 1

```{r}
library(knitr)
kable(summary(us_filtered))
```

This data is limited to 61 time points from June 2020 to July 2020. There are no missing points, since there are 30 days in June and 31 in July, so that gives a total of 61 days to survey.

#### Part 2

```{r}
library(stargazer)
#stargazer(fit)
```

\begin{table}[!htbp] \centering 
  \caption{} 
  \label{} 
\begin{tabular}{@{\extracolsep{5pt}}lc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{1}{c}{\textit{Dependent variable:}} \\ 
\cline{2-2} 
\\[-1.8ex] & `Cumulative\_number\_for\_14\_days\_of\_COVID-19\_cases\_per\_100000` \\ 
\hline \\[-1.8ex] 
 index & 4.107$^{***}$ \\ 
  & (0.145) \\ 
  & \\ 
 Constant & 42.853$^{***}$ \\ 
  & (5.165) \\ 
  & \\ 
\hline \\[-1.8ex] 
Observations & 61 \\ 
R$^{2}$ & 0.932 \\ 
Adjusted R$^{2}$ & 0.930 \\ 
Residual Std. Error & 19.922 (df = 59) \\ 
F Statistic & 803.464$^{***}$ (df = 1; 59) \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table}

### Part B

```{r}
#install.packages("broom")
fit.diags <- broom::augment(fit)
plot(fit,c(1:3,5))
```

### Part C

```{r}
acf(fit$residuals)
```

## Problem 5

```{r}
par(mfrow=c(2,2))
par(mar=c(2,2,2,2))
plot(fit, c(1:3,5))
```